---
title: "Fox & Weisberg "
output: html_document
---

```{r setup, include = F}

source("load_packages.R")

# install.packages("car", dependencies = T)
library(car)

# carPrestige #Candian occupation data
```

```{r}
data(Prestige)

#add column name to first column:
Prestige <- cbind(as_tibble(rownames(Prestige)), Prestige) %>% 
  rename(job = value) #can't wrap in as_tibble to drop row names as this messes up the labeling on Avplots 

Prestige
```

```{r}
prestige.mod.2 = lm(prestige ~ education + income + type, data = Prestige)
```

```{r}
#all predictors and overall
residualPlots(prestige.mod.2)

# no predictors, only overall
# residualPlots(prestige.mod.2, ~1) 

#only one predictor
# residualPlots(prestige.mod.2, ~education , fitted = F) 
```

Variation of basic residual plot above:
```{r}
#Univariate view:
marginalModelPlots(prestige.mod.2)

```

Added-variable plots (partial-regression plots). Partial relationship adjusted for all other variables in the model (as opposed to univariate above)
```{r}
avPlots(prestige.mod.2)#, id.n=2)#, id.cex=.6)
```

```{r}
#page 295
data(Duncan) #US occupation data
Duncan
```
```{r}
mod.duncan = lm(prestige ~ income + education, data = Duncan)
qqPlot(mod.duncan, id.n=3)
```
p 296
```{r}
outlierTest(mod.duncan)
```

hat-values
```{r}
suppressWarnings(influenceIndexPlot(mod.duncan , id.n=3))
```
```{r}
suppressWarnings(influencePlot(mod.duncan, id.n=3))
```

```{r}
mod.duncan.2 = update(mod.duncan, subset = rownames(Duncan) != "minister")
compareCoefs(mod.duncan, mod.duncan.2)
```

```{r}
mod.duncan.3 = update(mod.duncan, subset = !(rownames(Duncan) %in% c("conductor","minister")))
compareCoefs(mod.duncan, mod.duncan.2, mod.duncan.3, se=F)
```

p 300
```{r}
suppressWarnings(avPlots(mod.duncan, id.n = 3))
```
```{r}
dfbs_duncan = as_tibble(data.frame(rownames = row.names(dfbetas(mod.duncan)),dfbetas(mod.duncan)))
dfbs_duncan
```

```{r}

dfbs_duncan %>% 
  ggplot(aes(x = income, y = education))+
  geom_point()+
  geom_text(data = dfbs_duncan %>% filter(income < -.5 | income > .4),aes(income,education,label=rownames))


ggplotly(dfbs_duncan %>% 
  ggplot(aes(x = income, y = education))+
  geom_point(aes(text = rownames))
)
```


```{r}
#manual filter based on graph above to get rownames:
dfbs_duncan %>% 
  filter(education > 1)
```

p304
```{r}
mod.ornstein = lm(interlocks +1 ~ log(assets) + nation + sector, data = Ornstein)
# par(mfrow=c(1,2))
qqPlot(mod.ornstein, id.n=0)
plot(density(rstudent(mod.ornstein)))

```


```{r}
boxCox(mod.ornstein, lambda = seq(0,.6, by=.1))
#max lambda approx .22. 
```
similar to box Cox above but produces numeric instead of graphical output:
```{r}
p1 = powerTransform(mod.ornstein)
summary(p1)
```

```{r}
# Ornstein1_old = transform(Ornstein,
#                       y1=bcPower(interlocks +1, coef(p1)),
#                       y1round = bcPower(interlocks+1, coef(p1,round=T))
#                       ) #from the book

Ornstein1 = Ornstein %>% 
  mutate(#y1=bcPower(interlocks +1, coef(p1)), #not used. same results as rounded below
         interlocks_transform = bcPower(interlocks+1, lambda = coef(p1,round=T)) 
          )
Ornstein1
# summary(Ornstein1)

```

```{r}
mod.ornstein.trans = update(mod.ornstein, interlocks_transform ~., data = Ornstein1)

qqPlot(mod.ornstein.trans, id.n=0)
plot(density(rstudent(mod.ornstein.trans)))
```

constructed variable plot p 307
```{r}
mod.orstein.constr_var = update(mod.ornstein, .~. +boxCoxVariable(interlocks+1))
summary(mod.orstein.constr_var) #below is from book, but this is easier code and you just find that variable in table
# summary(mod.orstein.constr_var)$coef["boxCoxVariable(interlocks + 1)",]#drop = F]

```

```{r}

# avPlots(mod.orstein.constr_var) #have to hit enter in console but produces all vars
avPlots(mod.orstein.constr_var, variable = "boxCoxVariable(interlocks + 1)")


```

transforming predictors
compononet-plus-residual plots (or partial-residual-plots)
```{r}
prestige.mod.3 = update(prestige.mod.2, ~. - type + women)
crPlots(prestige.mod.3,order=2) #book says this is nearly identical to order=1 and CERES plots
crPlots(prestige.mod.3,order=1)
```

```{r}
prestige.mod.4 = update(prestige.mod.3, .~. +log2(income) - income)
crPlots(prestige.mod.4,order=2)
```

```{r}
prestige.mod.5 = update(prestige.mod.4, .~. - women + poly(women,2))
crPlots(prestige.mod.5,order=1)
summary(prestige.mod.4)$coef
summary(prestige.mod.5)$coef #why does poly have to coefs?
```
p 314
```{r}
residualPlots(mod.ornstein, ~1 ,fitted = T, id.n=0,quadratic=F,tests=F)
```
```{r}
spreadLevelPlot(mod.ornstein)
```

p 315 collinearity
```{r}
data(Ericksen)
Ericksen
```

```{r}
model_census = lm(undercount ~ . ,data=Ericksen)
summary(model_census)

model_census_2 = lm(undercount ~ . -highschool - poverty - housing - city,data=Ericksen)
summary(model_census_2)
```
Check for collinearity (vif > 4)
```{r}
vif(model_census) #can also be called on glms

```

```{r}
Ericksen %>% pull(undercount) %>% hist
model_census_glm = glm(undercount ~ . , data = Ericksen)
model_census_glm #same as lm() above since family not specified

```

```{r}
vif(model_census_glm)
```




