Hash MD5 on policy number and insured name for train/test/holdout split reproducibilty 
```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(openssl, stringr)

data = data.frame(pol_num = c('123-0','456-0'), insured_name = c('insured_1','insured_2'))

data = data %>% 
  mutate(hash = md5(str_c(pol_num,insured_name))) %>% 
  mutate(split = as.integer(as.hexmode(substr(hash, 1, 1)) %% 2L)) #modify 2L for different splits. what if not equal?
  #can also split by sorting on hash or cutoff values

data
```

Gini and Lorenz Curve

"To produce a Gini coefficient and a Lorenz curve plot for a model, it is necessary to modify the functions. Instead
of sorting based on actual data, one sorts by predicted values. It is also advisable to include a random number
column as a secondary sort key. If a model produces the same predicted value for large segments of the data, and
if those segments had been previously sorted in ascending order of actual values, and if the sort function preserves
that order, the Gini coefficient and Lorenz curve might look better than they ought to be. Introducing a random
number column as a secondary sort key avoids that issue."


re-written example from study note code 

```{r}
pacman::p_load(caTools)
set.seed(123) 

# chick_weight = as_tibble(ChickWeight) %>% 
#   mutate(count = 1) %>% #not sure this is necessary. used in weightedGini function that causes an error
#   mutate(sample = if_else(sample.split(weight, SplitRatio = .75) == T,"Train","Test")) #replace sample column above

#use same code as paper for consistent results
chick_weight = as_tibble(ChickWeight) %>% 
  mutate(count = 1) %>% 
  mutate(sample = as.factor(sample(c("Test","Train"), nrow(ChickWeight), replace=TRUE, prob = c(0.3,0.7))))


#check split close to target
nrow(chick_weight %>% filter(sample == "Train"))/nrow(chick_weight) 

chick_weight
```

Gini Coefficient and Lorenz Curve 
https://stackoverflow.com/questions/22679493/how-to-plot-a-nice-lorenz-curve-for-factors-in-r-ggplot
 ggplot(data=Distr1_df) +
    geom_point(aes(x=p, y=L)) +
    geom_line(aes(x=p, y=L), color="#990000") +
    scale_x_continuous(name="Cumulative share of X", limits=c(0,1)) + 
    scale_y_continuous(name="Cumulative share of Y", limits=c(0,1)) +
    geom_abline()
    
    library(ggplot2)
library(gglorenz)
    ggplot(x, aes(Distr1)) + 
  stat_lorenz() + 
  geom_abline(color = "grey")

```{r}
pacman::p_load(ineq,reldist)

train = chick_weight %>% filter(sample == "Train")

# quick output
gini(train$weight)
plot(Lc(train$weight))

#formatted output:
xtitle <- paste("Gini Coefficient =", round(gini(train$weight),5))
plot(Lc(train$weight), col = "blue", lwd = 2,
 main = "Training Sample", xlab = xtitle)
# round(with(train, WeightedGini(weight, count, weight)), 5) # email sent to CAS. no WeightedGini function


test = chick_weight %>% filter(sample == "Test")

xtitle <- paste("Gini Coefficient =", round(gini(test$weight),5))
plot(Lc(test$weight), col = "blue", lwd = 2,
 main = "Testing Sample", xlab = xtitle)
```
https://www.kaggle.com/c/liberty-mutual-fire-peril/discussion/9880
http://blog.nguyenvq.com/blog/2015/09/25/calculate-the-weighted-gini-coefficient-or-auc-in-r/
```{r}


WeightedGini <- function(actual, weights, predicted){
 # Modification of code posted in Kaggle by William Cukierski
 # https://www.kaggle.com/c/liberty-mutual-fire-peril/discussion/9880
 # actual = actual frequency, severity, loss cost
 # corresponding weights = exposure, claim count, exposure
 # predicted = predicted
 df = data.frame(actual = actual, weights = weights, predicted =
predicted)
 # create random number sort key so ties will be resolved in random order
 k <- length(df$actual)
 df$rkey <- runif(k)
 df <- df[order(df$predicted, df$rkey),]
 df$random = cumsum((df$weights/sum(df$weights)))
 totalPositive <- sum(df$actual * df$weights)
 df$cumPosFound <- cumsum(df$actual * df$weights)
 df$Lorentz <- df$cumPosFound / totalPositive
 n <- nrow(df)
 gini <- sum(df$Lorentz[-1]*df$random[-n]) - sum(df$Lorentz[-n] * df$random[-1])
 return(gini)
}

# Lorenz Curve function that uses predicted values for sorting
LorenzCurve2 <- function (x, n = rep(1, length(x)), xhat, plot = FALSE)
{
 # Modification of Lc function in ineq package
 # x = actual (frequency, severity, loss cost)
 # xhat = predicted (frequency, severity, loss cost)
 # n = weights (exposure, claims, exposure)
 # data will be sorted in ascending order of prediction
 # Lorenz Curve will be plotted based on actual values
 ina <- !is.na(x)
 n <- n[ina]
 x <- as.numeric(x)[ina]
 xhat <- as.numeric(xhat)[ina]
 k <- length(x)
 # create random number sort key so ties will be resolved in random order
 #rkey <- runif(k)
 #o <- order(xhat, rkey)
 o <- order(xhat)
 x <- x[o]
 n <- n[o]
 x <- n * x
 p <- cumsum(n)/sum(n)
 L <- cumsum(x)/sum(x)
 p <- c(0, p)
 L <- c(0, L)
 L2 <- L * mean(x)/mean(n)
 Lc <- list(p, L, L2)
 names(Lc) <- c("p", "L", "L.general")
 class(Lc) <- "Lc"
 if (plot)
 plot(Lc)
 Lc
}
```

```{r}
round(with(train, WeightedGini(weight, count, weight)), 5)
```

```{r}
pacman::p_load(randomForest)

set.seed(456)
model_rf = randomForest(weight ~ Diet + Time, data = train)

train_augment = train %>% 
  mutate(weight_p = predict(model_rf, train))

# debugonce(WeightedGini)
train_pgin = round(with(train_augment, WeightedGini(weight, count, weight_p)), 5)
train_pgin

xtitle = paste("Random Forest Gini Coefficient =",
 round(train_pgin,5))

with(train_augment, plot(LorenzCurve2(weight, count, weight_p),
 col="blue", lwd=2,
 main="Training Sample", xlab=xtitle))

# slightly different gini than the paper. train % same as source code. newer version of random forest?


# Predicted on Test Sample
test_augment = test %>% 
  mutate(weight_p = predict(model_rf, test))

test_pgin <- round(with(test_augment,
 WeightedGini(weight, count, weight_p)), 5)

xtitle <- paste("Random Forest Gini Coefficient =",
 round(test_pgin,5))

with(test_augment, plot(LorenzCurve2(weight, count, weight_p),
 col="blue", lwd=2,
 main="Test Sample", xlab=xtitle))

norm_gini_train = round(train_pgin/gini(train$weight),5)
norm_gini_test = round(test_pgin,5)/round(gini(test$weight),5)
message('Normalized Gini on training sample = ',round(train_pgin,5),"/",round(gini(train$weight),5)," = ",norm_gini_train)
message('Normalized Gini on testing sample = ',round(test_pgin,5),"/",round(gini(test$weight),5)," = ",norm_gini_test)
message('Difference in normalized Gini = ',norm_gini_train - norm_gini_test) #or cat() but then need to insert page breaks


```

#code yet to be recreated:
```{r}


#
# Fit linear regression training data
#

model_ols = lm(weight ~ Time,
 data = train)
# Predicted on Training Sample

train_augment = train %>% 
  mutate(weight_p = predict(model_ols, train))

train_pgin <- round(with(train_augment,
 WeightedGini(weight, count, weight_p)), 5)

xtitle <- paste("Linear Model Gini Coefficient =",
 round(train_pgin,5))

with(train_augment, plot(LorenzCurve2(weight, count, weight_p),
 col="blue", lwd=2,
main="Training Sample", xlab=xtitle))
# Predicted on Test Sample
test_augment = test %>% 
  mutate(weight_p = predict(model_ols, test))

test_pgin <- round(with(test_augment,
 WeightedGini(weight, count, weight_p)), 5)
xtitle <- paste("Linear Model Gini Coefficient =",
 round(test_pgin,5))
with(test_augment, plot(LorenzCurve2(weight, count, weight_p),
 col="blue", lwd=2,
 main="Test Sample", xlab=xtitle))

norm_gini_train = round(train_pgin/gini(train$weight),5)
norm_gini_test = round(test_pgin,5)/round(gini(test$weight),5)
message('Normalized Gini on training sample = ',round(train_pgin,5),"/",round(gini(train$weight),5)," = ",norm_gini_train)
message('Normalized Gini on testing sample = ',round(test_pgin,5),"/",round(gini(test$weight),5)," = ",norm_gini_test)
message('Difference in normalized Gini = ',norm_gini_train - norm_gini_test) #or cat() but then need to insert page breaks

```
PICKUP ON LIFT CHARTS

